Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
Clearing unneeded comments
Some corrections added

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

@@

\title{Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

% [Pedro] no aparecía el primer autor. En cualquier caso, lo dejo comentado a continuación.

%  \author{Juan-J.~Merelo, %Add your name here
%  \thanks{JJ Merelo is with the Dept. of Computer Architecture and Technology and CITIC University of Granada
%  e-mail: (jmerelo@ugr.es).}
%  }% <-this % stops a space

%  \author{Mario Garcia-Valdez, %Add your name here
%    \thanks{M Garcia-Valdez is with the Dept. of Graduate Studies at Instituto Tecnologico de Tijuana
%  e-mail: (mario@tectijuana.edu.mx).}
%  }% <-this % stops a space

@@

There is, then, space for systems that do not have any of those
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client, the latter  can be possibly
run on a  % [Pedro] Both the server and the client in browsers?
          % different browsers as I understand?  No, server in
          % server, client in browser, although there's nothing that
          % can't be done in a browser. 
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which
the scientist has no control.

Some time ago, our group faced  the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
hardware and continuous hardware failures of existing
clusters. Considering this, we leveraged our experience of working
with web applications, including JavaScript, since their inception and other volunteer and
unconventional distributed evolutionary computing systems, and we 
decided to designd and relase a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

Recent works propose the use of cloud computing services for the components of 
client-server distributed EAs. Cloud resources 
can reduce operational costs by outsourcing both hardware and software maintenance 
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what 
is available in most laboratories, allowing researches to 
scale algorithms at reasonable costs. Researchers have proposed the use of 
public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
and DropBox~\cite{mericloud}. The use of JavaScript to implement server side components, %FERGU: he puesto mejor el de revista en lugar del del CEC
gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen
in most Platforms as a Service (PaaS) as many support the Node.js framework \cite{wood13:nodejs:paas}. Also many web services %FERGU: he validado esta frase con un footnote, pero si quieres lo paso a referencias
have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
developers can use the same software tools for both client and server side development. 
Besides the advantages to deal with the same language in the full stack as previously mentioned, 
this also facilitates the interoperability of the exchanged data types.
%Ayuda con la mini conclusion FERGU: mira la última frase de arriba

In the work presented in this paper, JavaScript was used in the full 
stack, which has the advantage of using the same code for writing
desktop or command line and browser clients, both based in the EC
library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time when writing code and leverages %FERGU: párrafo en pasado "was used" y presente "saves"
expertise acquired in the library itself. %FERGU: tienes info de LoC usadas para la sección experimental y justificar esta frase?
%The proposed software has the advantage of full stack JavaScript
%development..
%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
Javascript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in 
metaheuristics optimization allows an improvement in communicability, 
reproducibility, interoperability, automated assembly, knowledge engineering 
and efficiency.

There have been relatively few efforts to analyze the
performance that can be obtained from these volunteer computing
efforts. 
%Too many efforts?
% No, too little - JJ 
% Digo efforts en la oración anterior :) - Mario
There were some initial attempts  to dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

@@

% [Pedro] desde el abstract no se había nombrado "NodIO". 
% Creo que en este párrafo debería introducirse, principalmente el nombre.
% Añado una frase presentándolo:
In this sense, in this paper we propose {\sf NodIO}, a PaaS based 
volunteer computing system, whose architecture
has been developed using JavaScript on both the client and server sides.
All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} has two parts:\begin{enumerate}
\item A REST server, that is, a server that includes several remote
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage or
  browser tab with the experiment is not currently on focus.  
\end{enumerate}

>>>>

\usepackage{array}

@@

\title{Volunteers in the Clouds: an Architecture for Low Cost and
  Potentially Massive Distributed Evolutionary Computation}

@@

There is, then, space for systems that do not have any of those
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client, the latter  can be possibly
run on a  % [Pedro] Both the server and the client in browsers?
          % different browsers as I understand? [JJ]  No, server in
          % server, client in browser, although there's nothing that
          % can't be done in a browser. 
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which
the scientist has no control.

Some time ago, our group faced  the problem of diminishing funds for
buying new     % [Pedro] This sentence is so so long. I would split it
               % in two sentences. [JJ] Done.
hardware and continuous hardware failures of existing
clusters. Considering this, we leveraged our experience of working
with web applications, including JavaScript, since their inception and other volunteer and
unconventional distributed evolutionary computing systems, and we 
decided to designd and relase a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

Recent works propose the use of cloud computing services for the components of 
client-server distributed EAs. Cloud resources 
can reduce operational costs by outsourcing both hardware and software maintenance 
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what 
is available in most laboratories, allowing researches to 
scale algorithms at reasonable costs. Researchers have proposed the use of 
public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
and DropBox~\cite{mericloud}. The use of JavaScript to implement server side components, %FERGU: he puesto mejor el de revista en lugar del del CEC
gives software developers certain advantages in the cloud paradigm. JavaScript is a first-classe citizen
in most Platforms as a Service (PaaS) as many support the Node.js framework \cite{wood13:nodejs:paas}. Also many web services %FERGU: he validado esta frase con un footnote, pero si quieres lo paso a referencias
have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
developers can use the same software tools for both client and server side development. 
Besides the advantages to deal with the same language in the full stack as previously mentioned, 
this also facilitates the interoperability of the exchanged data types.
%Ayuda con la mini conclusion FERGU: mira la última frase de arriba

In the work presented in this paper, JavaScript has been used throughout the full 
stack, which has the advantage of using the same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time
when writing code and leverages %FERGU: párrafo en pasado "was used" y
                                %presente "saves" 
% done - JJ
expertise acquired in the library itself. %FERGU: tienes info de LoC
                                %usadas para la sección experimental y
                                %justificar esta frase?
% Es el mismo código - JJ
%The proposed software has the advantage of full stack JavaScript
%development..
%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
Javascript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}. 
According to \cite{swanresearch2015}, the benefits of using this kind of paradigm in 
metaheuristics optimization allows an improvement in communicability, 
reproducibility, interoperability, automated assembly, knowledge engineering 
and efficiency.

There have been relatively few efforts to analyze the
performance that can be obtained from these volunteer computing
experiments. 
%Too many efforts?
% No, too little - JJ 
% Digo efforts en la oración anterior :) - Mario
% Visto :-) Gracias- JJ
There were some initial attempts to simply dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

@@

% [Pedro] desde el abstract no se había nombrado "NodIO". 
% Creo que en este párrafo debería introducirse, principalmente el nombre.
% Añado una frase presentándolo:
In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
volunteer computing system, whose architecture
has been developed using JavaScript on both the client and server sides.
All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} has two parts:\begin{enumerate}
\item A REST server, that is, a server that includes several remote
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage or
  browser tab with the experiment is not currently on focus.  
\end{enumerate}
